---
output:
  html_document: default
  pdf_document: default
---

```{r}
library(haven)
library(ggplot2)
data <- read_dta("D:/BrownUnivercity/DATA2020/Data2020-Final-Project/data/ff_data_x_preprocessed_v1.dta")
print(data)
```

**EDA**
1-2.cf1edu: Father baseline education; cm1edu: Mother baseline education; 1-low, 4-high

```{r}
# Filter the data to exclude rows where cf1edu equals -3
filtered_data <- data[data$cf1edu != -3, ]

# Create a bar plot for Father's education
ggplot(filtered_data, aes(x = cf1edu, fill = )) +
  geom_bar(alpha=0.9) +
  scale_fill_gradient(low='#483D8B', high='#DB7093') +
  labs(x="father education",title = "Father's Education")


```



```{r}
# Filter the data to exclude rows where cf1edu equals -3
filtered_data <- data[data$cf1edu != -3 & data$cm1edu != -3, ]

# Combine cf1edu and cm1edu into a single variable
combined_edu <- c(filtered_data$cf1edu, filtered_data$cm1edu)

# Create a corresponding group variable
group <- rep(c("Father's Education", "Mother's Education"), each = nrow(filtered_data))

# Combine data into a new dataframe
combined_data <- data.frame(combined_edu = factor(combined_edu), group = group)

# Create a bar plot for combined education levels
bar_plot <- ggplot(combined_data) +
  geom_bar(aes(x = combined_edu, fill = group),  position = "dodge") +
  labs(x = "Education Level", title = "Parents' Education") +
  scale_fill_manual(values = c("Father's Education" = "#483D8B", "Mother's Education" = "#FFD700"))
bar_plot
# ggsave(filename = "parents.png", plot = bar_plot)
```



3.f2b32: In general, how is your child's health?

```{r}
library(viridis)
# Filter the data without negative
filtered_data_h <- data[data$f2b32 > 0, ]
# boxplot
box_c <- ggplot(filtered_data_h, aes(x = y_binary, y = f2b32, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d()
ggsave(filename = "success by child's health1.png",plot = box_c)
# Violin Plot
filtered_data <- data[data$f2b32 > 0, ]
# violin plot
violin_c <- ggplot(filtered_data, aes(y = f2b32, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs( y = "child's healthe", x = "success", fill = "success") +
  theme_minimal()
ggsave(filename = "success by child's health2.png",plot=violin_c)
# histogram
filtered_data_m <- data[data$f2b32 > 0, ]
# Create a bar plot
bar_c <- ggplot(filtered_data_m, aes(x = f2b32, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') 
ggsave(filename = "success by child's health3.png",plot = bar_c)
```
4. m2d2:  Int Chk: Does father have any contact with child ?

```{r}
# Filter the data without negative
filtered_data_h <- data[data$m2d2 > 0, ]

# Boxplot
boxplot_plot <- ggplot(filtered_data_h, aes(x = y_binary, y = m2d2, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d() 

ggsave(filename = "father_contacts_with_child_boxplot.png", plot = boxplot_plot)

# Violin Plot
filtered_data <- data[data$f2b32 > 0, ]

violin_plot <- ggplot(filtered_data, aes(y = m2d2, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(y = "contact", x = "success", fill = "success") +
  theme_minimal()

ggsave(filename = "father_contacts_with_child_violin.png", plot = violin_plot)

# Histogram
filtered_data_m <- data[data$f2b32 > 0, ]

histogram_plot <- ggplot(filtered_data_m, aes(x = m2d2, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') 

ggsave(filename = "father_contacts_with_child_histogram.png", plot = histogram_plot)

```
5.m2b43a: On a scale of 1-(least like) to 5-(most like) - Child tends to be shy

6.cf3marm: Constructed - Is father married to child's mother at year three? 

```{r}
# 0-No 1-Yes
# Filter the data without negative  
filtered_data <- data[data$cf3marm >= 0, ]
# Count the frequency of each category in f2d1a
category_counts <- table(filtered_data$cf3marm)
# Convert the frequency table to a data frame
category_df <- as.data.frame(category_counts)
# Create a pie chart
mm <- ggplot(category_df, aes(x = "", y = Freq, fill = factor(Var1))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0)+
  theme_void() +
  theme(legend.position = "right")
# ggsave(filename = "father married to child's mother.png",plot=mm)
mm
```



```{r}
# Filter the data without negative values
filtered_data <- data[data$cf3marm >= 0, ]
# Count the frequency of each category in f2d1a
category_counts <- table(filtered_data$cf3marm)
# Convert the frequency table to a data frame
category_df <- as.data.frame(category_counts)
# Create a pie chart
mm <- ggplot(category_df, aes(x = "", y = Freq, fill = factor(Var1, labels = c("No", "Yes")))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "right") +
  labs(title = "Father married to child's mother at year three")
ggsave(filename = "father married to child's mother.png",plot=mm)
```


7.cf3kids: Constructed - Number of children under 18 in household
```{r}
# Filter the data without negative
filtered_data_h <- data[data$cf3kids > 0, ]

# Boxplot
boxplot_plot <- ggplot(filtered_data_h, aes(x = y_binary, y = cf3kids, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d() +
  labs(x="success",y="children under 18 in household")

ggsave(filename = "children_under_18_in_household_boxplot.png", plot = boxplot_plot)

# Violin Plot
violin_plot <- ggplot(filtered_data, aes(y = cf3kids, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(y = "cf3kids", x = "y_binary", fill = "y_binary") +
  theme_minimal()

ggsave(filename = "children_under_18_in_household_violin.png", plot = violin_plot)

# Histogram
histogram_plot <- ggplot(filtered_data_m, aes(x = cf3kids, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') 

ggsave(filename = "children_under_18_in_household_histogram.png", plot = histogram_plot)

```

8. cm3md_case_lib: Constructed - Father meets depression criteria (liberal) at three-year (CIDI)
```{r}
# Filter the data without negative  
filtered_data <- data[data$cf3md_case_lib >= 0, ]
# Count the frequency of each category in f2d1a
category_counts <- table(filtered_data$cf3md_case_lib)
# Convert the frequency table to a data frame
category_df <- as.data.frame(category_counts)
# Create a pie chart
nn <- ggplot(category_df, aes(x = "", y = Freq, fill = factor(Var1))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0)  +
  theme_void() +
  theme(legend.position = "right")
ggsave(filename = "Father meets depression criteria.png",plot=nn)
```
9.cf4cohm:  Constructed - Father living with child's mother at five-year

```{r}
# Filter the data without negative  
filtered_data <- data[data$cf3md_case_lib >= 0, ]
# Count the frequency of each category in f2d1a
category_counts <- table(filtered_data$cf3md_case_lib)
# Convert the frequency table to a data frame
category_df <- as.data.frame(category_counts)

bin <- ggplot(category_df, aes(x = "", y = Freq, fill = factor(Var1))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "right") +
  labs(title="Father living with child's mother at five-year",fill = "Success")+
  scale_fill_manual(values = c("0" = "#483D8B", "1" = "#FFD700"), labels = c("Not Success", "Success")) +
  scale_x_discrete(labels = c("0" = "Not Success", "1" = "Success"))
bin
# ggsave(filename = "Father living with child's mother at five-year.png",plot=bin)
```
10. t4d7: number of kids present with child

```{r}
filtered_data_m <- data[data$t4d7 >= 0, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = t4d7, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093')  +
  labs(x="kids present with child",title="Number of kids present with child")
ggsave(filename = "kids present with child.png")
```
11. cf1hhinc:  Household income 
```{r}
# Filter the data without negative
filtered_data <- data[data$cf1hhinc > 0, ]

# Boxplot
boxplot_plot <- ggplot(filtered_data, aes(x = y_binary, y = cf1hhinc, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.6)  +
  scale_fill_viridis_d() + 
  labs(x="success",y="Household income", title="Household income")

# ggsave(filename = "Household_income_boxplot.png", plot = boxplot_plot)

# Violin plot
violin_plot <- ggplot(filtered_data, aes(y = cf1hhinc, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs( y = "m2d2", x = "y_binary", fill = "y_binary") +
  theme_minimal()

# ggsave(filename = "Household_income_violin.png", plot = violin_plot)

# Histogram
histogram_plot <- ggplot(filtered_data, aes(x = cf1hhinc, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') 

# ggsave(filename = "Household_income_histogram.png", plot = histogram_plot)

boxplot_plot

```


```{r}
# Filter the data without negative
filtered_data <- data[data$cf1hhinc > 0, ]
boxplot_plot <- ggplot(filtered_data, aes(x = factor(y_binary), y = cf1hhinc, fill =as.factor(y_binary)))     +
  geom_boxplot(alpha = 0.5)  +
  scale_fill_manual(values = c("0" = "#483D8B", "1" = "#FFD700"), labels = c("Not Success", "Success")) +
  scale_x_discrete(labels = c("0" = "Not Success", "1" = "Success")) +
  labs(x = "Success", y = "Household income", title = "Household Income",fill="Success")
boxplot_plot
ggsave(filename = "Household_income_boxplot.png", plot = boxplot_plot)
```






12. k5d1f: Amount of time on a weekday you play computer games on the computer or TV.

```{r}
# Filter the data without negative
filtered_data <- data[data$k5d1f > 0, ]
# Mosaic Plot
colors <- c("#5F9EA0", "#ffca3a", "#6495ED","#6a4c93")
mm <- mosaicplot(table(filtered_data$y_binary, filtered_data$k5d1f),
           ylab = "Success Score",
           color = colors,
           main = "Success Score vs. Time on Computer or TV on a Weekday",
           sub = paste("0-none, 1-half an hour or less,",'\n',
            "2-more than half an hour but less than an hour,",'\n',
            "3-one to two hour, 4 more than 2 hours"))
ggsave(filename = "play computer games2.png",plot=mm)
```
13. k5f1f    F1F. Hurt an animal on purpose

```{r}
filtered_data_m <- data[data$k5f1f >= 0, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = k5f1f)) +
  geom_bar(fill='#DB7093')  +
  labs(title = "Hurt an animal on purpose")
ggsave(filename = "Hurt an animal on purpose.png")
# Mosaic Plot
colors <- c("#5F9EA0", "#ffca3a", "#6495ED","#6a4c93")
pp <- mosaicplot(table(filtered_data_m$y_binary, filtered_data_m$k5f1f),
           main = "Success Score vs. Hurt an animal on purpose ",
           ylab = "Success Score",
           color = colors,
           sub = paste("1-Yes,2-No"))
ggsave(filename = "Hurt an animal on purpose2.png",plot = pp)
```
```{r}
# Y 
library(haven)
library(ggplot2)
old_data <- read_dta("D:/BrownUnivercity/DATA2020/Data2020-Final-Project/data/ff_data_x_preprocessed.dta")

ggplot(old_data, aes(x = y_score, fill = ..x..)) +
  geom_histogram(bins = 7, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093')  
ggsave(filename = "y_old.png")
```


```{r}
# Y 
data <- read_dta("D:/BrownUnivercity/DATA2020/Data2020-Final-Project/data/ff_data_x_preprocessed_v1.dta")

ggplot(data, aes(x = y_binary, fill = ..x..)) +
  geom_histogram(bins = 2, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') 
ggsave(filename = "y_new.png")
```
```{r}
old_data <- read_dta("D:/BrownUnivercity/DATA2020/Data2020-Final-Project/data/ff_data_x_preprocessed.dta")

ggplot(old_data, aes(x = y_score, fill = factor(y_score))) +
  geom_bar() +
  scale_fill_manual(values = c("0" = "#595091", "1" = "#6F5992", "2" = "#836094", "3" = "#986997", "4" = "#A96F99", "5" = "#BD7999", "6" = "#CF819B")) +
  labs(x = "Success Score")
ggsave(filename = "y_old.png")
```
```{r}
library(dplyr)
data <- read_dta("D:/BrownUnivercity/DATA2020/Data2020-Final-Project/data/ff_data_x_preprocessed_v1.dta")

data$y_binary =  factor(data$y_binary, levels = c("1", "0"))
category_counts <- data %>% 
    group_by(y_binary) %>% 
    tally() %>% 
    arrange(desc(n))


# Create a stacked bar chart
ggplot(category_counts, aes(x = "", y = n, fill = y_binary)) +
    geom_bar(stat = "identity",width = 0.6) +
    labs(x = "Success Score",y = "Count", title = "Stacked Bar Chart of Target Variable Distribution",fill = "Sucess Score") +
    scale_fill_manual(values = c(
         
        "1" = "#3E0751","0" = "#F9E855"),labels = c( "Success","Not Success"
    ))+
  scale_x_discrete(labels = c( "1" = "Success","0" = "Not Success")) +
    coord_flip() 


```




















