---
output:
  html_document: default
  pdf_document: default
---

```{r}
library(haven)
library(ggplot2)
data <- read_dta("D:/BrownUnivercity/DATA2020/Data2020-Final-Project/data/ff_data_x_preprocessed_v1.dta")
print(data)
```

**EDA**
1-2.cf1edu: Father baseline education; cm1edu: Mother baseline education; 1-low, 4-high

```{r}
# Filter the data to exclude rows where cf1edu equals -3
filtered_data <- data[data$cf1edu != -3, ]
# Create a bar plot
ggplot(filtered_data, aes(x = cf1edu, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#808000', high='#FF7F50') + 
  labs(title = "Distribution of Father education")
ggsave(filename = "father.png")

# Mother
# Filter the data to exclude rows where cf1edu equals -3
filtered_data_m <- data[data$cm1edu != -3, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = cm1edu, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') + 
  labs(title = "Distribution of Mother education")
ggsave(filename = "mother.png")
```

3.f2b32: In general, how is your child's health?

```{r}
library(viridis)
# Filter the data without negative
filtered_data_h <- data[data$f2b32 > 0, ]
# boxplot
ggplot(filtered_data_h, aes(x = y_binary, y = f2b32, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d() +
  labs(title = "Distribution of success by child's health")
ggsave(filename = "success by child's health1.png")
# Violin Plot
filtered_data <- data[data$f2b32 > 0, ]
# violin plot
ggplot(filtered_data, aes(y = f2b32, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(title = "Violin Plot of success by child's health", y = "f2b32", x = "y_binary", fill = "y_binary") +
  theme_minimal()
ggsave(filename = "success by child's health2.png")
# histogram
filtered_data_m <- data[data$f2b32 > 0, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = f2b32, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') + 
  labs(title = "Distribution of success by child's health")
ggsave(filename = "success by child's health3.png")
```
4. m2d2:  Int Chk: Does father have any contact with child ?

```{r}
# Filter the data without negative
filtered_data_h <- data[data$m2d2 > 0, ]
# boxplot
ggplot(filtered_data_h, aes(x = y_binary, y = m2d2, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d() +
  labs(title = "Distribution of success by father contacts with child")
ggsave(filename = "father contacts with child.png")
# Violin Plot
filtered_data <- data[data$f2b32 > 0, ]
# violin plot
ggplot(filtered_data, aes(y = m2d2, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(title = "Violin Plot of Distribution of success by father contacts with child", y = "m2d2", x = "y_binary", fill = "y_binary") +
  theme_minimal()
ggsave(filename = "father contacts with child2.png")
# histogram
filtered_data_m <- data[data$f2b32 > 0, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = m2d2, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') + 
  labs(title = "Distribution of success by father contacts with child")
ggsave(filename = "father contacts with child3.png")
```
5.m2b43a: On a scale of 1-(least like) to 5-(most like) - Child tends to be shy

6.cf3marm: Constructed - Is father married to child's mother at year three? 

```{r}
# 0-No 1-Yes
# Filter the data without negative  
filtered_data <- data[data$cf3marm >= 0, ]
# Count the frequency of each category in f2d1a
category_counts <- table(filtered_data$cf3marm)
# Convert the frequency table to a data frame
category_df <- as.data.frame(category_counts)
# Create a pie chart
ggplot(category_df, aes(x = "", y = Freq, fill = factor(Var1))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Distribution of father married to child's mother at year three") +
  theme_void() +
  theme(legend.position = "right")
ggsave(filename = "father married to child's mother.png")
```

7.cf3kids: Constructed - Number of children under 18 in household
```{r}
# Filter the data without negative
filtered_data_h <- data[data$cf3kids > 0, ]
# boxplot
ggplot(filtered_data_h, aes(x = y_binary, y = cf3kids, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d() +
  labs(title = "Distribution of success by children under 18 in household")
ggsave(filename = "children under 18 in household.png")

# Violin Plot
filtered_data <- data[data$cf3kids > 0, ]
# violin plot
ggplot(filtered_data, aes(y = cf3kids, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(title = "Violin Plot of Distribution of success by children under 18 in household", y = "cf3kids", x = "y_binary", fill = "y_binary") +
  theme_minimal()
ggsave(filename = "children under 18 in household2.png")

# histogram
filtered_data_m <- data[data$cf3kids > 0, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = cf3kids, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') + 
  labs(title = "Distribution of success by children under 18 in household")
ggsave(filename = "children under 18 in household3.png")
```

8. cm3md_case_lib: Constructed - Father meets depression criteria (liberal) at three-year (CIDI)
```{r}
# Filter the data without negative  
filtered_data <- data[data$cf3md_case_lib >= 0, ]
# Count the frequency of each category in f2d1a
category_counts <- table(filtered_data$cf3md_case_lib)
# Convert the frequency table to a data frame
category_df <- as.data.frame(category_counts)
# Create a pie chart
ggplot(category_df, aes(x = "", y = Freq, fill = factor(Var1))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Distribution of Father meets depression criteria") +
  theme_void() +
  theme(legend.position = "right")
ggsave(filename = "Father meets depression criteria.png")
```
9.cf4cohm:  Constructed - Father living with child's mother at five-year

```{r}
# Filter the data without negative  
filtered_data <- data[data$cf4cohm >= 0, ]
# Count the frequency of each category in f2d1a
category_counts <- table(filtered_data$cf4cohm)
# Convert the frequency table to a data frame
category_df <- as.data.frame(category_counts)
# Create a pie chart
ggplot(category_df, aes(x = "", y = Freq, fill = factor(Var1))) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Father living with child's mother at five-year") +
  theme_void() +
  theme(legend.position = "right")
ggsave(filename = "Father living with child's mother at five-year.png")
```
10. t4d7: number of kids present with child

```{r}
filtered_data_m <- data[data$t4d7 >= 0, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = t4d7, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') + 
  labs(title = "Distribution of kids present with child")  
ggsave(filename = "kids present with child.png")
# scatter plot
# ggplot(filtered_data_m, aes(x = t4d7, y = y_binary)) +
  # geom_point() +
  # labs(title = "Success by kids present with child")
```
11. cf1hhinc:  Household income 
```{r}
# Filter the data without negative
filtered_data <- data[data$cf1hhinc > 0, ]
# boxplot
ggplot(filtered_data, aes(x = y_binary, y = cf1hhinc, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d() +
  labs(title = "Distribution of success by Household income")
ggsave(filename = "Household income.png")
# violin plot
ggplot(filtered_data, aes(y = cf1hhinc, x = y_binary, fill = as.factor(y_binary))) +
  geom_violin() +
  scale_fill_viridis_d() +
  labs(title = "Distribution of success by Household income", y = "m2d2", x = "y_binary", fill = "y_binary") +
  theme_minimal()
ggsave(filename = "Household income2.png")

# histogram
ggplot(filtered_data, aes(x = cf1hhinc, fill = ..x..)) +
  geom_histogram(bins = 10, alpha=0.9)  +
  scale_fill_gradient(low='#483D8B', high='#DB7093') + 
  labs(title = "Distribution of success by Household income")
ggsave(filename = "Household income3.png")
```
12. k5d1f: Amount of time on a weekday you play computer games on the computer or TV.

```{r}
# Filter the data without negative
filtered_data <- data[data$k5d1f > 0, ]
# boxplot
ggplot(filtered_data, aes(x = y_binary, y = k5d1f, fill = as.factor(y_binary))) +
  geom_boxplot(alpha = 0.3)  +
  scale_fill_viridis_d() +
  labs(title = "Amount of time on a weekday kids play computer games on the computer or TV")
ggsave(filename = "play computer games.png")
# Mosaic Plot
colors <- c("#5F9EA0", "#ffca3a", "#6495ED","#6a4c93")
mosaicplot(table(filtered_data$y_binary, filtered_data$k5d1f),
           main = "Success Score vs. Time on Computer or TV on a Weekday",
           ylab = "Success Score",
           color = colors,
           sub = paste("0-none, 1-half an hour or less,",'\n',
            "2-more than half an hour but less than an hour,",'\n',
            "3-one to two hour, 4 more than 2 hours"))
ggsave(filename = "play computer games2.png")
```
13. k5f1f    F1F. Hurt an animal on purpose

```{r}
filtered_data_m <- data[data$k5f1f >= 0, ]
# Create a bar plot
ggplot(filtered_data_m, aes(x = k5f1f)) +
  geom_bar(fill='#DB7093')  +
  labs(title = "Hurt an animal on purpose")
ggsave(filename = "Hurt an animal on purpose.png")
# Mosaic Plot
colors <- c("#5F9EA0", "#ffca3a", "#6495ED","#6a4c93")
mosaicplot(table(filtered_data_m$y_binary, filtered_data_m$k5f1f),
           main = "Success Score vs. Hurt an animal on purpose ",
           ylab = "Success Score",
           color = colors,
           sub = paste("1-Yes,2-No"))
ggsave(filename = "Hurt an animal on purpose2.png")
```























